Faint Object Detection in Multi-Epoch Observations\\ via Catalog Data Fusion

### 2016-11-09 ms.tex
Alex's final mods to the appendix, with minor cleaning by TL

As we introduce the peak constraints, we need to first consider the impact
on the gradient. The constrained probability distribution
is
\begin{equation}
	dw_x = \exp\left(-\frac{\hh^T H^{-1} \hh}{2}\right) 
			|\det g| \delta^{(2)}[\hh]			
			\frac{d^2 \hh} {2\pi|H|^{1/2}}.
\end{equation}
After integrating over $d^2\hh$ we get the extremum weight
\begin{equation}
	w_x = \frac{|\det g|}{2\pi|H|^{1/2}} =\frac{x^2-y^2}{4 \pi R^2}.
\end{equation}
This will multiply the unconstrained probability for the density of extremal 
points of the random field, 
\begin{equation}
		dn_{\rm{}pk}=\left[\frac{(x^2-y^2)y}{2\pi R^2} e^{-y^2}\,dy\right]
		\exp\left[-\frac{x^2 + 2\gamma x z + z^2}{2(1-\gamma^2)}\right] 
			\frac{\,dx\ dz}{2\pi \sqrt{1-\gamma^2}}.
\end{equation}
For a peak both eigenvalues of the second derivative tensor must be negative. 
In the rotated coordinates $(x,y)$, this means that $x<0$, and $0<y<-x$.
We can easily integrate over the allowed range of $y$ next, yielding
\begin{equation}
	\int_0^{-x} dy\, y\/ (x^2-y^2)\, e^{-y^2} 
		=\frac{1}{2}(x^2-1+e^{-x^2}).
\end{equation}
We are left with
\begin{equation}
	dn_{\rm{}pk} = \frac{(-1 + x^2 +e^{-x^2})}{4\pi R^2}  
	\exp\left[-\frac{x^2 + 2 x z \gamma + z^2 }{2 (1-\gamma^2)}\right]
	\frac{\,dx\ dz}{2\pi \sqrt{1-\gamma^2}}.
\end{equation}
Let us introduce the function $B(x)$ as
\begin{equation}
	B(s,b) = \sqrt{\frac{\pi}{b}} \exp\left({\frac{s^2}{2b}}\right)
	\left[1+ {\rm erf}\left( \frac{s}{\sqrt{2b}}\right)\right].
\end{equation}
Evaluating the integral over $-\infty<x\leq 0$ in Mathematica, we obtain
\begin{equation}
 n_{\rm{}pk}(s,\gamma)= 
		%\exp\left(-\frac{s^2}{2\gamma^2}\right) 
		\frac{e^{-\frac{s^2}{2\gamma^2}}}{8\pi^2R^2}
		\Bigg[
			(1-\gamma^2)s +[s^2 - \gamma^2 (1 + s^2)] B(s,1)
			+B(s, 3- 2\gamma^2),
		\Bigg]
\end{equation}
with
\begin{equation}
	s = \frac{\gamma z}{\sqrt{1-\gamma^2}}.
\end{equation}

We need to evaluate the shape parameter $\gamma$. Assume that the window 
function applied to the random field is a Gaussian with a scale $a$,
\begin{equation}
	w(r) = \frac{1}{2\pi a^2} \exp\left(-\frac{r^2}{2 a^2}\right).
\end{equation}
Its Fourier transform is also a Gaussian,
\begin{equation}
	W(k) = \exp\left(-\frac{k^2a^2}{2}\right).
\end{equation}
We model the sky noise as a white noise with a flat spectrum. Thus the
correlations in the measured random field are determined by the window function,
i.e.,
\begin{equation}
	P(k) = A\,|W(k)|^2 = A \exp\left(-k^2a^2\right).
\end{equation}
With this power spectrum it is straightforward to compute the scale and the shape 
parameters as
\begin{equation}
	\gamma^2 = \frac{1}{2}, \qquad R^2 = \frac{a^2}{2}.
\end{equation}

>>>>

As we introduce the peak constraints, we need to first consider the impact
on the gradient. The constrained probability distribution
is
\begin{equation}
	dw_x = \exp\left(-\frac{\hh^T H^{-1} \hh}{2}\right) 
			|\det g| \delta^{(2)}[\hh]			
			\frac{d^2 \hh} {2\pi|H|^{1/2}}.
\end{equation}
After integrating over $d^2\hh$ we get the extremum weight
\begin{equation}
	w_x = \frac{|\det g|}{2\pi|H|^{1/2}} =\frac{x^2-y^2}{4 \pi R^2}.
\end{equation}
This will multiply the unconstrained probability for the density of extremal 
points of the random field, 
\begin{equation}
		dn_{\rm{}pk}=\left[\frac{(x^2-y^2)y}{2\pi R^2} e^{-y^2}\,dy\right]
		\exp\left[-\frac{x^2 + 2\gamma x z + z^2}{2(1-\gamma^2)}\right] 
			\frac{\,dx\ dz}{2\pi \sqrt{1-\gamma^2}}.
\end{equation}
For a peak both eigenvalues of the second derivative tensor must be negative. 
In the rotated coordinates $(x,y)$, this means that $x<0$, and $0<y<-x$.
We can easily integrate over the allowed range of $y$ next, yielding
\begin{equation}
	\int_0^{-x} dy\, y\/ (x^2-y^2)\, e^{-y^2} 
		=\frac{1}{2}(x^2-1+e^{-x^2}).
\end{equation}
We are left with
\begin{equation}
	dn_{\rm{}pk} = \frac{(-1 + x^2 +e^{-x^2})}{4\pi R^2}  
	\exp\left[-\frac{x^2 + 2 x z \gamma + z^2 }{2 (1-\gamma^2)}\right]
	\frac{\,dx\ dz}{2\pi \sqrt{1-\gamma^2}}.
\end{equation}
Let us introduce the function $B(x)$ as
\begin{equation}
	B(s,b) = \sqrt{\frac{\pi}{b}} \exp\left({\frac{s^2}{2b}}\right)
	\left[1+ {\rm erf}\left( \frac{s}{\sqrt{2b}}\right)\right].
\end{equation}
Evaluating the integral over $-\infty<x\leq 0$ in Mathematica, we obtain
\begin{equation}
 n_{\rm{}pk}(z)= 
		%\exp\left(-\frac{s^2}{2\gamma^2}\right) 
		\frac{e^{-\frac{s^2}{2\gamma^2}}}{8\pi^2R^2}
		\Bigg[
			(1-\gamma^2)s +[s^2 - \gamma^2 (1 + s^2)] B(s,1)
			+B(s, 3- 2\gamma^2),
		\Bigg],
\label{eq:npk}
\end{equation}
with
\begin{equation}
	s = \frac{\gamma z}{\sqrt{1-\gamma^2}}.
\end{equation}
We get the full surface density of noise peaks, $\lambda$, by integrating the conditional surface density in Eq.~\ref{eq:npk} over all peak heights $z$:
\begin{equation}
	\lambda = \int_{-\infty}^{\infty} n_{pk}(z) dz.
\end{equation}
Finally, we need to evaluate the shape parameter $\gamma$. Assume that the window 
function applied to the random field is a Gaussian with a scale $a$,
\begin{equation}
	w(r) = \frac{1}{2\pi a^2} \exp\left(-\frac{r^2}{2 a^2}\right).
\end{equation}
Its Fourier transform is also a Gaussian,
\begin{equation}
	W(k) = \exp\left(-\frac{k^2a^2}{2}\right).
\end{equation}
We model the sky noise as a white noise with a flat spectrum. Thus the
correlations in the measured random field are determined by the window function,
i.e.,
\begin{equation}
	P(k) = A\,|W(k)|^2 = A \exp\left(-k^2a^2\right).
\end{equation}
With this power spectrum it is straightforward to compute the scale and the shape parameters as
\begin{equation}
	\gamma^2 = \frac{1}{2}, \qquad R^2 = \frac{a^2}{2}.
\end{equation}
With this choice of PSF and $\gamma$, we get $s=z$ in Eq.(A25).

Now we are in a position to compute the probability that a noise peak is within a radius $r$ of our point of interest located at the origin. The spatial distribution of the noise peaks is described by a Poisson process with the surface density $\lambda$. 
The cumulative probability that the peak is within a radius $r$ is given by the well-known expression
\begin{equation}
	P_{pk}(<r|z) = 1-\exp[-\lambda(z) r^2 \pi].
\end{equation}
The differential probability is given by its derivative with respect to to $r$, as 
\begin{equation}
 p_{pk}(r|z) = 2\pi r \lambda(z) \exp[-\lambda(z) r^2 \pi]
\end{equation}
Both of these probabilities are shown on Fig.~\ref{fig:pk-shift}. The differential probability starts off around the origin scaling with $r$, due to the available area (phase space for configuration).
This in turn causes the cumulative function to rise as $r^2$, resulting in a very small probability ($<0.03$) that a noise peak will appear within a PSF scale.
Thus we can safely ignore noise peaks as a major contributor to false detections at a significant level.

\begin{figure}
\epsscale{0.95}
\plottwo{fig/app-fig-a3.pdf}{fig/app-fig-a4.pdf}
\caption{The differential and cumulative probabilities of having a noise peak at or within a radius $r$, integrated over all heights. The radius is shown in units of $a$, the PSF scale.
}
\label{fig:pk-shift}
\end{figure}
